### 2016-05-18 Paper_v1.tex
May17

% Load packages
\usepackage{url}  % Formatting web addresses
\usepackage{ifthen}  % Conditional
\usepackage{multicol}   %Columns
\usepackage[utf8]{inputenc} %unicode support
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{graphicx}
\usepackage[margin=0.1pt,font=footnotesize,labelfont=bf]{caption}
\usepackage{setspace}
%\usepackage{longtable}
\usepackage{colortbl}
%\usepackage{palatino,lettrine}
%\usepackage{times}
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage[wide]{sidecap}
%\usepackage[authoryear,round,comma,sort&compress]{natbib}
\usepackage[square,sort,comma,numbers,sort&compress]{natbib}
%\usepackage[authoryear,round]{natbib}
\usepackage{supertabular}
\usepackage{simplemargins}
\usepackage{fullpage}
\usepackage{comment}
\usepackage{lineno}
%\usepackage{chicago}
\usepackage{textcomp}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
\DeclareMathOperator*{\argmin}{\arg\!\min}

\usepackage{algorithm2e}
\usepackage{algpseudocode}
%\usepackage[space]{cite}
\urlstyle{rm}

@@

\subsection*{Formulation and Solution of the Model Equations}
We used ordinary differential equations (ODEs) to model the time evolution of metabolite ($x_{i}$) and scaled enzyme abundance ($\epsilon_{i}$) in hypothetical cell-free metabolic networks:
\begin{eqnarray}
	\frac{dx_{i}}{dt} & = & \sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right)\qquad{i=1,2,\hdots,\mathcal{M}}\\
	\frac{d\epsilon_{i}}{dt} & = & -\lambda_{i}\epsilon_{i}\qquad{i = 1,2,\hdots,\mathcal{E}}
\end{eqnarray}where $\mathcal{R}$ denotes the number of reactions, $\mathcal{M}$ denotes the number of metabolites and $\mathcal{E}$ denotes the number of enzymes in the model.
The quantity $r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right)$ denotes the rate of reaction $j$.
Typically, reaction $j$ is a non-linear function of metabolite and enzyme abundance, as well as unknown kinetic parameters $\mathbf{k}$ ($\mathcal{K}\times{1}$).
The quantity $\sigma_{ij}$ denotes the stoichiometric coefficient for species $i$ in reaction $j$.
If $\sigma_{ij}>0$, metabolite $i$ is produced by reaction $j$.
Conversely, if $\sigma_{ij}<0$, metabolite $i$ is consumed by reaction $j$, while $\sigma_{ij} = 0$ indicates metabolite $i$ is not connected with reaction $j$.
Lastly, $\lambda_{i}$ denotes the scaled enzyme degradation constant.
The system material balances were subject to the initial conditions $\mathbf{x}\left(t_{o}\right) = \mathbf{x}_{o}$ and $\mathbf{\epsilon}\left(t_{o}\right) = \mathbf{1}$ (initially we have 100\% cell-free enzyme abundance).

The reaction rate was written as the product of a kinetic term ($\bar{r}_{j}$) and a control term ($v_{j}$), $r_{j}\left(\mathbf{x},\mathbf{k}\right) = \bar{r}_{j}v_{j}$.
In this study, we used either saturation or mass action kinetics.
The control term $0\leq v_{j}\leq 1$ depended upon the combination of factors which influenced rate process $j$.
For each rate, we used a rule-based approach to select from competing control factors.
If rate j was influenced by $1,\dots,m$ factors, we modeled this relationship as
$v_{j} = \mathcal{I}_{j}\left(f_{1j}\left(\cdot\right),\hdots,f_{mj}\left(\cdot\right)\right)$
where $0\leq f_{ij}\left(\cdot\right)\leq 1$ denotes a regulatory transfer function quantifying the influence of factor $i$ on rate $j$.
The function $\mathcal{I}_{j}\left(\cdot\right)$ is an integration rule which maps the output of regulatory transfer functions into a control
variable. Each regulatory transfer function took the form:
\begin{equation}\label{eqn:control-factor}
	f_{ij}\left(\mathcal{Z}_{i},k_{ij},\eta_{ij}\right) = k_{ij}^{\eta_{ij}}\mathcal{Z}_{i}^{\eta_{ij}}/\left({1 + k_{ij}^{\eta_{ij}}\mathcal{Z}_{i}^{\eta_{ij}}}\right)
\end{equation}where $\mathcal{Z}_{i}$ denotes the abundance factor $i$, $k_{ij}$ denotes a gain parameter, and $\eta_{ij}$ denotes a cooperativity parameter.
In this study, we used $\mathcal{I}_{j}\in\left\{\max,\min\right\}$ \cite{pr3010138}. If a process has no modifying factors, $v_{j} = 1$.
We used multiple saturation kinetics to model the reaction term $\bar{r}_{j}$:
\begin{equation}\label{eqn:rate-bar}
	\bar{r}_{j} = k_{j}^{max}\epsilon_{i}\left(\prod_{s\in{m_{j}^{-}}}\frac{x_{s}}{K_{js} + x_{s}}\right)
\end{equation}
where $k_{j}^{max}$ denotes the maximum rate for reaction $j$, $\epsilon_{i}$ denotes the scaled enzyme activity which catalyzes reaction $j$, and
$K_{js}$ denotes the saturation constant for species $s$ in reaction $j$.
The product in Equation~\eqref{eqn:rate-bar} was carried out over the set of \textit{reactants} for reaction $j$ (denoted as $m_{j}^{-}$).

\subsection*{Generation of model ensemble}
Fill me in.

\subsection*{Global and local sensitivity analysis}
We conducted a global sensitivity analysis, using the variance-based method of Sobol, to estimate which parameters controlled the performance of the reduced order model \citep{SOBOL_METHOD}.
We computed the total sensitivity index of each parameter relative to two performance objectives, the peak thrombin time and the area under the thrombin curve (thrombin exposure).
We established the sampling bounds for each parameter from the minimum and maximum value of that parameter in the parameter set ensemble.
We used the sampling method of Saltelli \textit{et al.} \citep{Saltelli:2010} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
where $N$ was the number of trials, and $d$ was the number of parameters in the model. In our case, $N$ = 10,000 and $d$ = 22, so the total sensitivity indices were computed from
460,000 model evaluations. The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \citep{SALIB}.

@@

% Figures and captions go here ...
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.00\textwidth]{./figs/<Filename>.pdf}
%\caption{Captiontext goes here}
%}\label{fig:<label_name>}
%\end{figure}

>>>>

% Load packages
\usepackage{url}  % Formatting web addresses
\usepackage{ifthen}  % Conditional
\usepackage{multicol}   %Columns
\usepackage[utf8]{inputenc} %unicode support
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{graphicx}
\usepackage[margin=0.1pt,font=footnotesize,labelfont=bf]{caption}
\usepackage{setspace}
%\usepackage{longtable}
\usepackage{colortbl}
%\usepackage{palatino,lettrine}
%\usepackage{times}
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage[wide]{sidecap}
%\usepackage[authoryear,round,comma,sort&compress]{natbib}
\usepackage[square,sort,comma,numbers,sort&compress]{natbib}
%\usepackage[authoryear,round]{natbib}
\usepackage{supertabular}
\usepackage{simplemargins}
\usepackage{fullpage}
\usepackage{comment}
\usepackage{lineno}
%\usepackage{chicago}
\usepackage{textcomp}
\usepackage{multirow}
\usepackage{amsmath}
%\usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
\DeclareMathOperator*{\argmin}{\arg\!\min}

%\usepackage{algorithm2e}
%\usepackage{algpseudocode}
%\usepackage[space]{cite}
\urlstyle{rm}

@@

\subsection*{Formulation and Solution of the Model Equations}
We used ordinary differential equations (ODEs) to model the time evolution of metabolite ($x_{i}$) and scaled enzyme abundance ($\epsilon_{i}$) in hypothetical cell-free metabolic networks:
\begin{eqnarray}
	\frac{dx_{i}}{dt} & = & \sum_{j=1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right)\qquad{i=1,2,\hdots,\mathcal{M}}\\
	\frac{d\epsilon_{i}}{dt} & = & -\lambda_{i}\epsilon_{i}\qquad{i=1,2,\hdots,\mathcal{E}}
\end{eqnarray}where $\mathcal{R}$ denotes the number of reactions, $\mathcal{M}$ denotes the number of metabolites and $\mathcal{E}$ denotes the number of enzymes in the model.
The quantity $r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right)$ denotes the rate of reaction $j$.
Typically, reaction $j$ is a non-linear function of metabolite and enzyme abundance, as well as unknown kinetic parameters $\mathbf{k}$ ($\mathcal{K}\times{1}$).
The quantity $\sigma_{ij}$ denotes the stoichiometric coefficient for species $i$ in reaction $j$.
If $\sigma_{ij}>0$, metabolite $i$ is produced by reaction $j$.
Conversely, if $\sigma_{ij}<0$, metabolite $i$ is consumed by reaction $j$, while $\sigma_{ij}=0$ indicates metabolite $i$ is not connected with reaction $j$.
Lastly, $\lambda_{i}$ denotes the scaled enzyme degradation constant.
The system material balances were subject to the initial conditions $\mathbf{x}\left(t_{o}\right)=\mathbf{x}_{o}$ and $\mathbf{\epsilon}\left(t_{o}\right)=\mathbf{1}$ (initially we have 100\% cell-free enzyme abundance).

The reaction rate was written as the product of a kinetic term ($\bar{r}_{j}$) and a control term ($v_{j}$), $r_{j}\left(\mathbf{x},\mathbf{k}\right)=\bar{r}_{j}v_{j}$.
In this study, we used either saturation or mass action kinetics.
The control term $0\leq v_{j}\leq 1$ depended upon the combination of factors which influenced rate process $j$.
For each rate, we used a rule-based approach to select from competing control factors.
If rate j was influenced by $1,\dots,m$ factors, we modeled this relationship as
$v_{j}=\mathcal{I}_{j}\left(f_{1j}\left(\cdot\right),\hdots,f_{mj}\left(\cdot\right)\right)$
where $0\leq f_{ij}\left(\cdot\right)\leq 1$ denotes a regulatory transfer function quantifying the influence of factor $i$ on rate $j$.
The function $\mathcal{I}_{j}\left(\cdot\right)$ is an integration rule which maps the output of regulatory transfer functions into a control
variable. Each regulatory transfer function took the form:
\begin{equation}\label{eqn:control-factor}
	f_{ij}\left(\mathcal{Z}_{i},k_{ij},\eta_{ij}\right)=k_{ij}^{\eta_{ij}}\mathcal{Z}_{i}^{\eta_{ij}}/\left({1 + k_{ij}^{\eta_{ij}}\mathcal{Z}_{i}^{\eta_{ij}}}\right)
\end{equation}where $\mathcal{Z}_{i}$ denotes the abundance factor $i$, $k_{ij}$ denotes a gain parameter, and $\eta_{ij}$ denotes a cooperativity parameter.
In this study, we used $\mathcal{I}_{j}\in\left\{mean\right\}$ \cite{pr3010138}. If a process has no modifying factors, $v_{j}=1$.
We used multiple saturation kinetics to model the reaction term $\bar{r}_{j}$:
\begin{equation}\label{eqn:rate-bar}
	\bar{r}_{j}=k_{j}^{max}\epsilon_{i}\left(\prod_{s\in{m_{j}^{-}}}\frac{x_{s}}{K_{js} + x_{s}}\right)
\end{equation}
where $k_{j}^{max}$ denotes the maximum rate for reaction $j$, $\epsilon_{i}$ denotes the scaled enzyme activity which catalyzes reaction $j$, and
$K_{js}$ denotes the saturation constant for species $s$ in reaction $j$.
The product in Equation~\eqref{eqn:rate-bar} was carried out over the set of \textit{reactants} for reaction $j$ (denoted as $m_{j}^{-}$).

\subsection*{Generation of model ensemble}
We generated an ensemble of 18,000 parameter sets via a downhill-only random walk Monte Carlo method \cite{?}.
Beginning with a single parameter set as a starting point, we calculated its cost function, equal to the sum-absolute-error between experimental data and model predicitons:
\begin{equation}\label{eqn:cost-function}
    cost=\sum_{i=1}^{\mathcal{D}}\left(w_i\sum_{j=1}^{\mathcal{T}}abs\left(x_{ij}^{data}-x_{i}^{sim}|_{t(j)}\right)\right)
\end{equation}
where $\mathcal{D}$ denotes the number of datasets, $w_i$ denotes a weight, equal to 5 for the glucose, CAT, pyruvate, lactate, acetate, succinate, and malate datasets, and 1 elsewhere, $\mathcal{T}$ denotes the number of timepoints in the $i$th dataset, $t(j)$ denotes the $j$th timepoint, $x_{ij}^{data}$ denotes the value of the $i$th dataset at the $j$th timepoint, and $x_{i}^{sim}|_{t(j)}$ dneotes the simulated value of the metabolite corresponding to the $i$th dataset, interpolated to the $j$th timepoint.
We then perturbed model parameters:
\begin{equation}\label{eqn:parameter-perturbation}
    k_i^{new}=k_i*exp(a\,r_i)\qquad{i=1,2,\hdots,\mathcal{P}}
\end{equation}
where $\mathcal{P}$ denotes the number of parameters, equal to 652, which includes 163 rate constants, 455 saturation constants, and 34 control parameters, $k_i^{new}$ denotes the new value of the $i$th parameter, $k_i$ denotes the current value of the $i$th parameter, $a$ denotes a distribution variance, set to 0.03, and $r$ denotes a random sample from the normal distribution.
We stored the parameter set and calculated its cost; if it was less than the previous cost, we used the new parameter set to generate the following set.
After generating 180,000 sets we defined the 18,000 sets with the lowest cost values as our ensemble, and the set with the lowest cost value as our best-fit set.

\subsection*{Global and local sensitivity analysis}
We conducted a global sensitivity analysis, using the variance-based method of Sobol, to estimate which of the experimentally controllable parameters affected the performance of the reduced order model \citep{SOBOL_METHOD}.
This included the initial conditions of glucose, oxygen, amino acids, and enzymes.
We computed the total sensitivity index of each parameter relative to a performance objective of area under the CAT curve (CAT production).
We established the sampling bounds for each parameter from the value of that parameter in the set used to generate the ensemble.
We used the sampling method of Saltelli \textit{et al.} \citep{Saltelli:2010} to compute a family of $N\left(2d+2\right)$ sets which obeyed our parameter ranges,
where $N$ was the number of trials, and $d$ was the number of parameters in the model. In our case, $N$ = 300 and $d$ = 185, so the total sensitivity indices were computed from
111,600 model evaluations. The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \citep{SALIB}.
We conducted a local sensitivity analysis to estimate which of the other model parameters affected performance.
This included the same parameters that were varied in the ensemble: rate constants, saturation constants, and control parameters.
The local sensitivity for each parameter was calculated across a sub-ensemble of 180 parameter sets, randomly chosen from the ensemble of 18,000 sets:
\begin{equation}\label{eqn:local-sensitivity}
\begin{split}
    S_{ij}=\frac{p_{ij}}{AUC(p_{ij})}\frac{AUC(p_{ij}+\Delta p_{ij})-AUC(p_{ij})}{\Delta p_{ij}}\qquad{i=1,2,\hdots,\mathcal{E}}\qquad{j=1,2,\hdots,\mathcal{P}} \\
    \Delta p_{ij}=0.001\;p_{ij}\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad
\end{split}
\end{equation}
where $\mathcal{E}$ denotes the number of parameter sets in the sub-ensemble, equal to 180, $\mathcal{P}$ denotes the number of parameters, equal to 652, $S_{ij}$ denotes the sensitivity of the $j$th parameter for the $i$th parameter set, $p_{ij}$ denotes the value of the $j$th parameter for the $i$th parameter set, $\Delta p_{ij}$ denotes the perturbation of the $j$th parameter for the $i$th parameter set, equal to 0.1\% of the parameter value, and $AUC()$ denotes the area under the CAT curve.
We then calculated the mean and

\subsection*{Calculation of CAT yield}
The theoritical carbon yield of CAT was calculated using flux balance anaylsis (FBA) with a sequence-based analysis on CAT.
The sequence specific FBA \cite{2002_allen_palsson} problem was formulated as:
\begin{equation}\label{eqn:FBA-yield}
\begin{split}
	\max_{\boldsymbol{v}}{} \! \left( v_{obj}=\mathbf{\boldsymbol{c}}^T \boldsymbol{v} \right) \qquad \alpha_i \leq v_i \leq \beta_i  \qquad i=1,2,\hdots,\mathcal{R} \\
	\mathrm{Subject \; to:}	\; \; \mathbf{S}\mathbf{v}=\mathbf{0}\qquad\qquad\qquad\qquad\qquad\qquad\qquad\quad
\end{split}
\end{equation}
where $\mathbf{S}$ denotes the stoichiometric matrix, $\mathbf{v}$ denotes the unknown flux vector, $\boldsymbol{c}$ denotes the objective selection vector, and $\alpha_i$ and $\beta_i$ denote the lower and upper bounds on flux $v_{i}$, respectively.
The stoichiometric matrix was expanded to include the transcription and translation reactions for producing CAT. 
The objective $v_{obj}$ was to maximize the specific rate of CAT formation.
The specific glucose, amino acids and oxygen uptake rates were constrained to allow a maximum flux of 10 mM/hr.
The flux balance analysis problem was solved using the GNU Linear Programming Kit (v4.52) \cite{GLPK}.
The solution flux vector was used to calculate the theoritical carbon yield of CAT based on carbon consumed (glucose and amino acids).
The carbon yield was formulated as:
\begin{equation}\nonumber
	Yield=\frac{C_{CAT}v_{CAT}}{\sum_{i=1}^{\mathcal{R}}C_{i}v_{i}}
\end{equation}
where $C_{CAT}$ and $C_{i}$ denote the carbon number of CAT and substrate $i$, respectively, $v_{CAT}$ and $v_{i}$ denote the flux of CAT and substrate $i$, respectively,
and $\mathcal{R}$ denotes the number of substrates consumed.

@@

% Figures and captions go here ...
\begin{figure}[ht]
\centering
\includegraphics[width=0.75\textwidth]{./Figures/Network.pdf}
\caption{Flux profile for glycolysis, pentose phosphate pathway, Entner-Doudoroff pathway, TCA cycle, and NADPH/NADH transfer. Mean ± standard error for the reaction flux at 1.5 hrs (top) and 3 hrs (bottom), normalized to CAT synthesis flux.}
\label{fig:Network}
\end{figure}

\begin{figure}[ht]
\centering
%\includegraphics[width=1.00\textwidth]{./Figures/BothCarbon.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/Carbon.pdf}
\includegraphics[width=0.05\textwidth]{./Figures/null.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/CarbonNoControl.pdf}
\caption{A: Central carbon metabolism in the presence of allosteric control, including glucose (substrate), CAT (product), and intermediates, as well as total concentration of energy species. Best-fit parameter set (orange line) versus experimental data (points). 95\% confidence interval (blue shaded region) and 95\% confidence interval of the mean (gray shaded region) over the ensemble of 18,000 sets. B: Central carbon metabolism in the absence of allosteric control.}
\label{fig:BothCarbon}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/Energy.pdf}
\caption{Energy species and energy totals by base in the presence of allosteric control. Best-fit parameter set (orange line) versus experimental data (points). 95\% confidence interval (blue shaded region) and 95\% confidence interval of the mean (gray shaded region) over the ensemble of 18,000 sets.}
\label{fig:Energy}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/Amino.pdf}
\caption{Amino acids in the presence of allosteric control. Best-fit parameter set (orange line) versus experimental data (points). 95\% confidence interval (blue shaded region) and 95\% confidence interval of the mean (gray shaded region) over the ensemble of 18,000 sets.}
\label{fig:Amino}
\end{figure}

\begin{figure}[ht]
\centering
%\includegraphics[width=1.00\textwidth]{./Figures/SensitivitiesHalfTall.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/Sensitivity.pdf}
\includegraphics[width=1.00\textwidth]{./Figures/Metab_sensitivity.pdf}
\caption{Total-order global sensitivities for experimentally controllable initial conditions, including glucose, oxygen, amino acids, and enzymes.}
\label{fig:SensitivitiesHalfTall}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=0.94\textwidth]{./Figures/SensAll.png}
\caption{Mean and standard error of local sensitivities of rate constants (top), saturation constants (middle), and control parameters (bottom).}
\label{fig:SensAll}
\end{figure}
